### 2023-03-23 resilience.tex
details
up
Up

The resilience problem (resp. the $k$-resilience problem) in a transition system is to decide whether from a bad state, there exists a path (resp. a path of length smaller than $k$) that reaches a \Safe state. Let us formalize these properties.

@@

\alain{trouver les \Bad et \Safe maximum tels que S est resilient. est-ce vrai que si S est $(B_i,D_i)$-resilient alors S est $(\cap, \cup B_i,D_i)$-resilient ? on pourrait décider que \Safe=complement de \Bad et donc on aurait un seul ensemble (et son complement)}

@@

\begin{theorem}\label{up-up}
The % three %uniform 
resilience problem is decidable for WSTS with effective predbasis, $\Safe=\uparrow \Safe$
and $\Bad=\uparrow \Bad$.
%	and \Bad is upward closed or downward closed.
\end{theorem}

\begin{proof}
Let us first solve if $\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
(\Bad,\Safe)-resilient. Since $\Safe=\uparrow \Safe$, and WSTS, $\pred^*(\Safe)=\uparrow \pred^*(\Safe)$, so $\pred^*(\Safe)$ admits a finite basis $B_{\pred^*(\Safe)}$. since WSTS  with effective predbasis, we may compute a finite basis of $\pred^*(\Safe)$ with the backward coverability algorithm. 
%Then if \Bad is upward closed or downward closed, we may decide the inclusion $\Bad \subseteq \uparrow B_{\pred^*(\Safe)}$ which remais to compare, in both cases, two finite sets. 
if \Bad  is upward closed, $\Bad = \uparrow B_{\Bad}$ and $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$. 
\iffalse
similar reasonning for the other case.
\fi
hence the %two
 %uniform
 resilience problem is decidable.
\end{proof}

@@

\begin{theorem}\label{up-up}
The % three %uniform 
$k$-resilience problem is decidable for WSTS with effective predbasis, $\Safe=\uparrow \Safe$
and $\Bad=\uparrow \Bad$.
%	and \Bad is upward closed or downward closed.
\end{theorem}

\begin{proof}
Let us examine the %uniform 
$k$-resilience problem (kRP). We may compute the least $n$ such that $\Bad \subseteq \uparrow \pred^n(\Safe)=  \uparrow \pred^*(\Safe)$. $S$ is %uniformely 
$k$-resilient iff $n \leq k$ \alain{hum...mais il peut y avoir des chemins plus courts}
\mathieu{Si on a la résilience alors on a la $k$-résilience pour les $k$ plus grands que le $n$ minimal tel que $\uparrow \pred^n(\Safe)=  \uparrow \pred^*(\Safe)$, mais il faut traiter aussi les $k$ plus petits.}
\mathieu{Si on a $\uparrow \pred^k(\Safe) = \pred^k(\Safe)$ on résoud comme ci-dessus.}
\mathieu{... mais pour les $k$ plus petits, on peut avoir $\uparrow \pred^k(\Safe) \neq \pred^k(\Safe)$.}
\mathieu{... et alors, c'est plus compliqué.}
\mathieu{L'important c'est d'avoir le $n$ minimal tel que $\Bad \subseteq \pred^n(\Safe)$, qui n'est pas forcément tel que $\uparrow \pred^n(\Safe)=  \uparrow \pred^*(\Safe)$.}and then kRP is decidable.
\end{proof}

%One begins to test whether $\Bad \cap \pred^*(\Safe) = \emptyset$ then never resilient car une fois dans \Bad j'y reste sans pouvoir aller dans \Safe.\\

If $\Bad \cap \pred^*(\Safe) \neq \emptyset$, One computes the minimal elements $A$ of $\Bad \cap \pred^*(\Safe)$.
resilient iff $\Bad \subseteq \pred^*(\Safe)$  (eq $\Bad \cap \pred^*(\Safe) = \Bad$)

$\Bad \cap \pred^*(\Safe) \neq \Bad$ alors, on peut se limiter à $\Bad' = \Bad \cap \pred^*(\Safe)$ et c'est (\Bad',\Safe)-resilient.

Let us show that the predbasis hypothesis can be replaced by other hypotheses.

@@

\begin{proposition}\label{down-up}
Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and two finite sets: $I_{\Bad}$ and $Min_{\Safe}$.
The %uniform 
resilience problem (RP) and the %uniform 
bounded resilience problem (BRP) are decidable.
\end{proposition}

@@

To decide the %(uniform)
 bounded resilience problem (BRP), we decide $n \times m$ coverability questions: is state $b_j$ coverable from ideal $J_i$ ? If all these $n \times m$ coverability questions are positive then we compute $K=max(k_a \mid a=1,..,m \times n)$ and we deduce that  $\mathscr{S}$ is %uniformely
  $K$-resilient. Since coverability is decidable for completion-post-effective $\omega^2$-WSTS, we have shown that both the %uniform 
  resilience problem (BRP) and the %uniform 
  bounded resilience problem (BRP) are decidable.
\end{proof}

\begin{proposition}
Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and the predbasis hypothesis, and two finite sets: $I_{\Bad}$ and $Min_{\Safe}$.
The %uniform
 $k$-resilience problem (kRP) is decidable.
 \end{proposition}

@@

\begin{tabular}{ l c r }
\hline   \Bad/\Safe & Up & Down \\ \hline
   Up & dec ? & dec ? \\ \hline
   Down & dec & ? \\ \hline
 \end{tabular}

>>>>

The resilience problem (resp. the $k$-resilience problem) in a transition system is to decide whether from a bad state, there exists a path (resp. a path of length smaller than $k$) that reaches a safe state. Let us formalize these properties.

@@

\alain{trouver les \Bad~ et \Safe~ maximum tels que S est resilient. est-ce vrai que si S est $(B_i,D_i)$-resilient alors S est $(\cap, \cup B_i,D_i)$-resilient ? on pourrait décider que \Safe=complement de \Bad~ et donc on aurait un seul ensemble (et son complement)}

@@

\begin{theorem}\label{up-up}
The resilience problem is decidable for WSTS with effective predbasis, $\Safe=\uparrow \Safe$
and $\Bad=\uparrow \Bad$.

\end{theorem}

\begin{proof}
% Let us first solve if $\mathscr{S}=(S,\rightarrow,\leq)$ is (\Bad,\Safe)-resilient. 
Since $\Safe=\uparrow \Safe$ and
$\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS,  $\pred^*(\Safe)=\uparrow \pred^*(\Safe)$ and $\pred^*(\Safe)$ admits a finite basis $B_{\pred^*(\Safe)}$. Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective predbasis, we may compute a finite basis of $\pred^*(\Safe)$ with the backward coverability algorithm. 
Since $\Bad$  is upward closed, there exists a finite basis $B_{\Bad}$ such that $\Bad = \uparrow B_{\Bad}$. Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$,
hence the resilience problem is decidable.
\end{proof}

@@

\begin{theorem}\label{k-up-up}
The % three %uniform 
$k$-resilience problem is decidable for WSTS with effective predbasis, strong upward-compatibility, $\Safe=\uparrow \Safe$
and $\Bad=\uparrow \Bad$.
%	and \Bad is upward closed or downward closed.
\end{theorem}

\begin{proof}
Since $\Safe=\uparrow \Safe$ and
$\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong upward-compatibility,  $\pred^k(\Safe)=\uparrow \pred^k(\Safe)$ for all $k \in \N$, and $\pred^k(\Safe)$ admits a finite basis $B_{\pred^k(\Safe)}$.
\end{proof}

@@

\begin{theorem}\label{down-up}
Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and two finite sets: $I_{\Bad}$ and $Min_{\Safe}$.
The resilience problem (RP) and the bounded resilience problem (BRP) are decidable.
\end{theorem}

@@

To decide the  bounded resilience problem (BRP), we decide $n \times m$ coverability questions: is state $b_j$ coverable from ideal $J_i$ ? If all these $n \times m$ coverability questions are positive then we compute $K=max(k_a \mid a=1,..,m \times n)$ and we deduce that  $\mathscr{S}$ is   $K$-resilient. Since coverability is decidable for completion-post-effective $\omega^2$-WSTS, we have shown that both the 
  resilience problem (BRP) and the  
  bounded resilience problem (BRP) are decidable.
\end{proof}

\begin{theorem}
Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and the predbasis hypothesis, and two finite sets: $I_{\Bad}$ and $Min_{\Safe}$.
The  $k$-resilience problem (kRP) is decidable.
 \end{theorem}

@@

\begin{tabular}{ | l | c | c | c | r |}
\hline   \Safe~\Bad & $\uparrow$~ $\uparrow$~ & $\uparrow$~ $\downarrow$~ & $\downarrow$~ $\uparrow$~ & $\downarrow$~ $\downarrow$~ \\ \hline
   RP & Decidable (Theorem~\ref{up-up}) & Decidable (Theorem~\ref{down-up}) & ? Undecidable & Undecidable \\ \hline
   BRP & Decidable & ? Decidable (Theorem~\ref{down-up}) & ? & Undecidable \\ \hline
      kRP & Decidable (Theorem~\ref{k-up-up}) & Decidable & ? & Undecidable \\ \hline
 \end{tabular}
